\textbf{Department Invited Speakers Do Not Reflect Trainee Diversity}

### 2019-09-10 submission/manuscript.Rmd
work on results

Ada K. Hagan, Ph.D.${^1}$^, Rebecca M. Pollet, Ph.D.${^1}$^, and Josie Libertucci, Ph.D.${^2}$^

@@

Figures:

Tables:

@@

## Methods
Each academic year, each faculty member in the Department of Microbiology and Immunology at the University of Michigan has the opportunity to invite one speaker per year for a weekly seminar series. Some of these seminar slots are dedicated to named lectureships, which are decided by committee, and three trainee-invited speakers. We analyzed the gender and demographics of invited speakers and faculty hosts for five academic years (Fall 2014 - Spring 2019), and the current trainees when the data were analyzed (Spring 2019).

Each speaker was only counted once and any departmental faculty members or those listed as a "host" at any point could not also be considered "invited speakers". The list of faculty hosts was used as a proxy for faculty demographics since as hosts, these faculty members are visible representatives of the department. The list of trainees was obtained from department listservs and included masters students, doctoral students, and post-doctoral fellows.

We hand-coded demographics using personal knowledge, photos, and CVs. The presenting gender if each invidual was assigned using a binary system (man/woman). Diversity definitions vary according to the goals and population in question. However, in the United States, there is an inclination to consider both individuals of historically under-represented minority (HURM) and international backgrounds together. We argue that it is important to distinguish these two groups separately since each face different issues in the US and the academy and thus require different support systems. For instance, international scientists must contend with visa issues while HURMs have the trauma associated with living in a country who systematically shuts them out (despite an infrastructure that was built on their historical land and labor). For this reason, other assigned demographics included Caucasian, Historically Under-represented Minority (HURM), and International, each with a binary (yes/no) possibilty. Caucasian was assigned using the current U.S. Census definition where those of Middle Eastern, European, and Russian descent are included. HURM individuals were restricted to those with African-American, Indigenous and/or Hispanic heritage while International individuals were either visiting the US at the time of their talk, or immigrated to the US as an adult.

## Results and Discussion

To understand the representation of women, we compared the proportion of individuals presenting as women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to X% of faculty hosts and X% of the invited speakers. 10 of 31 lectureships [what about lectureships?] The proportion of women as faculty hosts and speakers is equivalent to global estimates of women microbiologists (Elseiver). Several papers have investigated the representation of women at scientific conferences, however, we only identified one that focused on invited speakers at university-sponsored events (Nittrouer, 2018). [Compare to our findings]

Assumptions about competency and dedication may negatively impact the likelihood of women microbiologists being invited as speakers. The dedication of women with family to their work is also more likely to be questioned than that of their colleagues, even men who also have children. For instance, the perceived prioritization and commitments of women to family over work may cause faculty to doubt their acceptance of a speaking invitation (despite the prestigious nature of these invitations), causing the faculty member to invite a different colleague who they feel is less likely to turn them down. Departments have different processes and criteria for selecting invited speakers, but want to bring the best scientists possible. It may be that the definition of "best" poses a problem to women, who need three-times as many publications as their men colleges to be considered equally competent. Some departments only invite tenured faculty, which severely limits the number of potential women speakers considerably. Alternately, pre-tenure faculty members invite prestigious, tenured faculty in their field to network and secure letters for their own tenure package. The increased burden of women to prove competency decreases their likelihood to be considered for either tenure and as possible source of tenure letters.

Our analysis identified an over-representation of Caucasian individuals as hosting faculty and invited speakers, relative to the proportion of Caucasian trainees. We observed declines in the representation of HURM and international faculty and speakers relative to the trainees.

[Lectureships]. 22, 3, 6, (cauc, HURM, Internatl, respectively) -- 12, 10, 5, 0 (CaucM, CaucW, Non-CaucM, Non-CaucW)

@@

Table 1. List of Suggestions & Resources

>>>>

Ada K. Hagan, Ph.D.${^1\dagger}$, Rebecca M. Pollet, Ph.D.${^1}$, and Josie Libertucci, Ph.D.${^2\dagger}$

@@

Figures: 1

Tables: 1

@@

## Methods
Each academic year, each faculty member in the Department of Microbiology and Immunology at the University of Michigan has the opportunity to invite one speaker per year for a weekly seminar series. Some of these seminar slots are dedicated to named lectureships, which are decided by committee, and three trainee-invited speakers. We analyzed the demographics of invited speakers and faculty hosts for five academic years (Fall 2014 - Spring 2019), and compared them to the current trainees when the data were analyzed (Spring 2019). Each speaker was only counted once and those listed as departmental faculty members or as a "host" at any point could not also be considered "invited speakers". The list of faculty hosts was used as a proxy for faculty demographics since as hosts, these faculty members are visible representatives of the department. The trainee lists were obtained from department listservs that included masters students, doctoral students, and post-doctoral fellows.

We hand-coded demographics using personal knowledge, photos, and CVs. The presenting gender of each individual was assigned using a binary system (man/woman). Diversity definitions vary according to the goals and population in question. However, in the United States, there is an inclination to consider both  together. We believe that it is important to distinguish between individuals of historically under-represented minority (HURM) and international backgrounds since each face different issues in the US and the academy and thus require different support systems. For instance, international scientists must contend with visa issues while HURMs have the trauma associated with living in a country who systematically shuts them out (despite an infrastructure that was built on their historical land and labor). For this reason, other assigned demographics included Caucasian, Historically Under-represented Minority (HURM), and International, each with a binary (yes/no) possibilty. Caucasian was assigned using the current U.S. Census definition where those of Middle Eastern, European, and Russian descent are included. HURM individuals were restricted to those with African-American, Indigenous and/or Hispanic heritage while International individuals were either visiting the US at the time of their seminar, or immigrated to the US as an adult.

## Results

To understand the representation of women, we compared the proportion of women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to `r gender_combined[[1,5]]`% of faculty hosts and `r gender_combined[[3,5]]`% of the invited speakers (Fig. 1A). Of 27 lectureships over the five year period, `r get_percent(10, 27)`% were awarded to women. The proportion of women as faculty hosts and speakers is equivalent to global estimates that 40% of microbiologists are women (Elseiver), with a slightly lower representation of women in lectureships.

Our analysis identified an over-representation of Caucasian individuals as hosting faculty and invited speakers, relative to the proportion of Caucasian trainees (Fig. 1B). We also observed declines in the representation of HURM and international faculty and speakers relative to the trainees, particularly postdocs (Fig 1B). Caucasian scientists also dominated lectureships, comprising `r get_percent(22, 27)`% of those awarded (Fig. 1C). Three and six lectureships were awarded to HURM and International scientists, respectively. Because the intersection of identities can compound biases and outcomes, we further examined the more prestigous lectureships by gender and Caucasian status. Caucasian men and women accounted for `r get_percent(12, 27)`% and `r get_percent(10, 27)`% of the lectureships, respectively, compared to `r get_percent(5, 27)`% non-Caucasian men and zero non-Caucasian women (Fig. 1D).

## Discussion

Several papers have investigated the representation of women at scientific conferences, however, we have only identified one that focused on invited speakers at universities (Nittrouer, 2018). In their study, Nittrouer et. al., examined 3,652 talks at 50 U.S. institutions in 2013 - 2014 and found that women faculty are less likely to be invited speakers, despite similar acceptance rates.

These results suggest that women faculty are less likely to invited as speakers, a decision that may be negatively impacted by assumptions about competency and dedication. The dedication of women who have children to their work is perceived to be less than that of their colleagues, e.g., men who also have children. The perceived prioritization and commitments of women to family over work may cause faculty to doubt their acceptance of a speaking invitation (despite the prestigious nature of these invitations and evidence to the contrary), causing the faculty member to invite a different colleague who they feel is more likely to agree. Departments have different processes and criteria for selecting invited speakers, but want to bring the best scientists possible. It may also be that the definition of "best" poses a problem to women, who need three-times as many publications as their men colleges to be considered equally competent. Some departments only invite tenured faculty, which severely limits the number of potential women speakers. Alternately, pre-tenure faculty members invite prestigious, tenured faculty in their field to network and secure letters for their own tenure package. The increased burden of women to prove competency decreases their likelihood to be considered for either tenure and as possible source of tenure letters.

+ studies that go beyond gender -- e.g., race/ethnicity (this seems to be the first)

@@

\begin{center}
\captionof{table}{List of suggestions and resources to increase invited speaker diversity.}
\small
\begin{tabular}{|l|l|l|}
\hline

\rowcolor{lightgray}
\textbf{Suggestion} & \textbf{Description} & \textbf{Resource} \\ \hline

Lab-invited speakers & \makecell[l]{Faculty members can request \\suggestions from trainees} & \\ \hline

Use a list & \makecell[l]{Many lists of scientists from \\under-represented and underserved \\groups are available} &  \makecell[l]{https://DiversifyMicrobiology.\\github.io/resources}\\ \hline

Create a list & \makecell[l]{Use the GitHub template \\ create a self-nomination list and \\resource for your field} & \makecell[l]{https://github.com/diversifymicrobiology/\\DiversifyMicrobiology.github.io} \\ \hline

Highlight the journey & \makecell[l]{Invite all speakers to spend \\a few moments describing their \\personal science journey} & \\ \hline

\end{tabular}
\end{center}

### 2019-09-10 submission/manuscript.tex
work on results

Ada K. Hagan, Ph.D.\({^1}\)\^{}, Rebecca M. Pollet, Ph.D.\({^1}\)\^{},
and Josie Libertucci, Ph.D.\({^2}\)\^{}

@@

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the gender and demographics of invited speakers and faculty
hosts for five academic years (Fall 2014 - Spring 2019), and the current
trainees when the data were analyzed (Spring 2019).

Each speaker was only counted once and any departmental faculty members
or those listed as a ``host'' at any point could not also be considered
``invited speakers''. The list of faculty hosts was used as a proxy for
faculty demographics since as hosts, these faculty members are visible
representatives of the department. The list of trainees was obtained
from department listservs and included masters students, doctoral
students, and post-doctoral fellows.

We hand-coded demographics using personal knowledge, photos, and CVs.
The presenting gender if each invidual was assigned using a binary
system (man/woman). Diversity definitions vary according to the goals
and population in question. However, in the United States, there is an
inclination to consider both individuals of historically
under-represented minority (HURM) and international backgrounds
together. We argue that it is important to distinguish these two groups
separately since each face different issues in the US and the academy
and thus require different support systems. For instance, international
scientists must contend with visa issues while HURMs have the trauma
associated with living in a country who systematically shuts them out
(despite an infrastructure that was built on their historical land and
labor). For this reason, other assigned demographics included Caucasian,
Historically Under-represented Minority (HURM), and International, each
with a binary (yes/no) possibilty. Caucasian was assigned using the
current U.S. Census definition where those of Middle Eastern, European,
and Russian descent are included. HURM individuals were restricted to
those with African-American, Indigenous and/or Hispanic heritage while
International individuals were either visiting the US at the time of
their talk, or immigrated to the US as an adult.

\subsection{Results and Discussion}\label{results-and-discussion}

To understand the representation of women, we compared the proportion of
individuals presenting as women in each academic role. At the trainee
level, more than half of students and postdoctoral fellows were women.
That dropped to X\% of faculty hosts and X\% of the invited speakers. 10
of 31 lectureships {[}what about lectureships?{]} The proportion of
women as faculty hosts and speakers is equivalent to global estimates of
women microbiologists (Elseiver). Several papers have investigated the
representation of women at scientific conferences, however, we only
identified one that focused on invited speakers at university-sponsored
events (Nittrouer, 2018). {[}Compare to our findings{]}

Assumptions about competency and dedication may negatively impact the
likelihood of women microbiologists being invited as speakers. The
dedication of women with family to their work is also more likely to be
questioned than that of their colleagues, even men who also have
children. For instance, the perceived prioritization and commitments of
women to family over work may cause faculty to doubt their acceptance of
a speaking invitation (despite the prestigious nature of these
invitations), causing the faculty member to invite a different colleague
who they feel is less likely to turn them down. Departments have
different processes and criteria for selecting invited speakers, but
want to bring the best scientists possible. It may be that the
definition of ``best'' poses a problem to women, who need three-times as
many publications as their men colleges to be considered equally
competent. Some departments only invite tenured faculty, which severely
limits the number of potential women speakers considerably. Alternately,
pre-tenure faculty members invite prestigious, tenured faculty in their
field to network and secure letters for their own tenure package. The
increased burden of women to prove competency decreases their likelihood
to be considered for either tenure and as possible source of tenure
letters.

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers, relative to the proportion of
Caucasian trainees. We observed declines in the representation of HURM
and international faculty and speakers relative to the trainees.

{[}Lectureships{]}. 22, 3, 6, (cauc, HURM, Internatl, respectively) --
12, 10, 5, 0 (CaucM, CaucW, Non-CaucM, Non-CaucW)

@@

We thank Harry Mobley and the Department of Microbiology \& Immunology,
University of Michigan for their support and input. We would also like
to acknowledge Nick Lesniak and Dr.~Ariangela Kozick for their comments
and suggestions.

@@

A.K.H. collected the data, assigned demographics, analyzed the data, and
created the website. R.P. created the Google lists, forms, and website
content and the description of their maintenance. J.L. wrote the
introduction and provided conceptual advice. All authors contributed to
the final manuscript.

@@

Table 1. List of Suggestions \& Resources

>>>>

Ada K. Hagan, Ph.D.\({^1\dagger}\), Rebecca M. Pollet, Ph.D.\({^1}\),
and Josie Libertucci, Ph.D.\({^2\dagger}\)

@@

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the demographics of invited speakers and faculty hosts for five
academic years (Fall 2014 - Spring 2019), and compared them to the
current trainees when the data were analyzed (Spring 2019). Each speaker
was only counted once and those listed as departmental faculty members
or as a ``host'' at any point could not also be considered ``invited
speakers''. The list of faculty hosts was used as a proxy for faculty
demographics since as hosts, these faculty members are visible
representatives of the department. The trainee lists were obtained from
department listservs that included masters students, doctoral students,
and post-doctoral fellows.

We hand-coded demographics using personal knowledge, photos, and CVs.
The presenting gender of each individual was assigned using a binary
system (man/woman). Diversity definitions vary according to the goals
and population in question. However, in the United States, there is an
inclination to consider both together. We believe that it is important
to distinguish between individuals of historically under-represented
minority (HURM) and international backgrounds since each face different
issues in the US and the academy and thus require different support
systems. For instance, international scientists must contend with visa
issues while HURMs have the trauma associated with living in a country
who systematically shuts them out (despite an infrastructure that was
built on their historical land and labor). For this reason, other
assigned demographics included Caucasian, Historically Under-represented
Minority (HURM), and International, each with a binary (yes/no)
possibilty. Caucasian was assigned using the current U.S. Census
definition where those of Middle Eastern, European, and Russian descent
are included. HURM individuals were restricted to those with
African-American, Indigenous and/or Hispanic heritage while
International individuals were either visiting the US at the time of
their seminar, or immigrated to the US as an adult.

\subsection{Results}\label{results}

To understand the representation of women, we compared the proportion of
women in each academic role. At the trainee level, more than half of
students and postdoctoral fellows were women. That dropped to 46.77\% of
faculty hosts and 38.73\% of the invited speakers (Fig. 1A). Of 27
lectureships over the five year period, 37.04\% were awarded to women.
The proportion of women as faculty hosts and speakers is equivalent to
global estimates that 40\% of microbiologists are women (Elseiver), with
a slightly lower representation of women in lectureships.

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers, relative to the proportion of
Caucasian trainees (Fig. 1B). We also observed declines in the
representation of HURM and international faculty and speakers relative
to the trainees, particularly postdocs (Fig 1B). Caucasian scientists
also dominated lectureships, comprising 81.48\% of those awarded (Fig.
1C). Three and six lectureships were awarded to HURM and International
scientists, respectively. Because the intersection of identities can
compound biases and outcomes, we further examined the more prestigous
lectureships by gender and Caucasian status. Caucasian men and women
accounted for 44.44\% and 37.04\% of the lectureships, respectively,
compared to 18.52\% non-Caucasian men and zero non-Caucasian women (Fig.
1D).

\subsection{Discussion}\label{discussion}

Several papers have investigated the representation of women at
scientific conferences, however, we have only identified one that
focused on invited speakers at universities (Nittrouer, 2018). In their
study, Nittrouer et. al., examined 3,652 talks at 50 U.S. institutions
in 2013 - 2014 and found that women faculty are less likely to be
invited speakers, despite similar acceptance rates.

These results suggest that women faculty are less likely to invited as
speakers, a decision that may be negatively impacted by assumptions
about competency and dedication. The dedication of women who have
children to their work is perceived to be less than that of their
colleagues, e.g., men who also have children. The perceived
prioritization and commitments of women to family over work may cause
faculty to doubt their acceptance of a speaking invitation (despite the
prestigious nature of these invitations and evidence to the contrary),
causing the faculty member to invite a different colleague who they feel
is more likely to agree. Departments have different processes and
criteria for selecting invited speakers, but want to bring the best
scientists possible. It may also be that the definition of ``best''
poses a problem to women, who need three-times as many publications as
their men colleges to be considered equally competent. Some departments
only invite tenured faculty, which severely limits the number of
potential women speakers. Alternately, pre-tenure faculty members invite
prestigious, tenured faculty in their field to network and secure
letters for their own tenure package. The increased burden of women to
prove competency decreases their likelihood to be considered for either
tenure and as possible source of tenure letters.

@@

We thank Dr.~Harry Mobley and the Department of Microbiology \&
Immunology, University of Michigan for their input and financial support
that enabled publication of our manuscript. We would also like to
acknowledge Nick Lesniak and Dr.~Ariangela Kozick for their comments and
suggestions.

@@

A.K.H. collected the data, assigned demographics, analyzed the data, and
created the website. R.M.P. created the Google lists, forms, and website
content and the description of their maintenance. J.L. wrote the
introduction and provided conceptual advice. All authors contributed to
the final manuscript.

@@

\begin{center}
\captionof{table}{List of suggestions and resources to increase invited speaker diversity.}
\small
\begin{tabular}{|l|l|l|}
\hline

\rowcolor{lightgray}
\textbf{Suggestion} & \textbf{Description} & \textbf{Resource} \\ \hline

Lab-invited speakers & \makecell[l]{Faculty members can request \\suggestions from trainees} & \\ \hline

Use a list & \makecell[l]{Many lists of scientists from \\under-represented and underserved \\groups are available} &  \makecell[l]{https://DiversifyMicrobiology.\\github.io/resources}\\ \hline

Create a list & \makecell[l]{Use the GitHub template \\ create a self-nomination list and \\resource for your field} & \makecell[l]{https://github.com/diversifymicrobiology/\\DiversifyMicrobiology.github.io} \\ \hline

Highlight the journey & \makecell[l]{Invite all speakers to spend \\a few moments describing their \\personal science journey} & \\ \hline

\end{tabular}
\end{center}
